\Large Analyzing the Supreme Court Citation Network

### 2017-12-28 Tex/paper.tex
finishing dependency effects

\textbf{Mutuality:} When deciding a set of related cases, the courts opinions regularly reference each other in order to reinforce their arguments...

@@

The city, supported by several amici, argues for affirmance nonetheless. It asserts first that the federal civil service statute is not just a federal retirement provision unrelated to the ADEA, but in fact establishes age as a BFOQ for federal firefighters based on factors that properly go into that determination under the ADEA, see Western Air Lines, Inc. v. Criswell, post p. 472 U. S. 400. Second, the city asserts, a congressional finding that age is a BFOQ for a certain occupation is dispositive of that determination with respect to nonfederal employees in that occupation. We consider each of these contentions in turn.

@@

\textbf{Also popularity} If j served as precedent for i, j is authoritative regarding the rule of i, and should also be cited. Citations beget citations. Citation confers authority. Cases are even discussed in opinions in terms of the extent of influence they have had on other cases,

@@

\textbf{Sociality} Citations sent beget citations sent.  For each case discussed, often related cases to discuss, even those that do not cite each other. For each case that applies, there is likely a case that needs to be discussed in terms of why it does not apply. Justices often clarify not only those legal principles that apply, but often those that do not. When multiple lines of legal reasoning are drawn upon to justify a decision, often more are discussed in terms of their lack of applicability.

>>>>

\textbf{Reciprocity:} Reciprocity (also referred to as mutuality) is the tendency for node B to send a tie to node A in response to or coordination with A sending a tie to B [CITE FROM SYLLABUS]. It is typically not possible for reciprocated ties to form in legal opinions. Most citations reference past opinions that were issued before the citing opinion's case was even argued before the Court.  However, opinions written within the same Supreme Court term are often drafted in tandem, and can cite each other reciprocally.  Opinion A citing opinion B within the same term represents a signal that opinion B is relevant to the legal reasoning underpinning opinion A. Unlike the citations themselves, the applicability of legal rules or lines of reasoning across cases is not directed---if A is relevant to B, B is highly likely to be relevant to A. We expect opinions written within the same term to exhibit a high degree of reciprocity. Below we provide two example passages from opinions that illustrate the phenomenon of within-term reciprocity.

@@

The city, supported by several amici, argues for affirmance nonetheless. It asserts first that the federal civil service statute is not just a federal retirement provision unrelated to the ADEA, but in fact establishes age as a BFOQ for federal firefighters based on factors that properly go into that determination under the ADEA, see Western Air Lines, Inc. v. Criswell, post p. 472 U. S. 400. Second, the city asserts, a congressional finding that age is a BFOQ for a certain occupation is dispositive of that determination with respect to nonfederal employees in that occupation.

@@

\textbf{Popularity} Popularity, also termed ``preferential attachment'' is the tendency for ties to be sent to nodes to which many ties have already been sent [CITE FROM SYLLABUS]. Citations to an opinion signal both the Court's awareness of the legal reasoning of the case and the Court's evaluation that the opinion is an authoritative precedent. The more citations, the stronger this signal. Landmark cases, or those that establish new legal rules, are particularly authoritative and accrue citations from most future opinions that follow the respective line of reasoning. The passage below, from XXXX, in which an authoritative opinion is referenced, and even discussed in terms of the number of other cases by which it was followed. We consider each of these contentions in turn. Based on this popularity phenomenon, we expect that new citations will be disproportionately directed at cases to which many past citations have accrued.

@@

\textbf{Sociality} Sociality is the tendency for ties sent to beget more ties sent [CITE]. Structurally, this means that there are few nodes that send a moderate number of ties---the number of ties sent is either small or fairly large. In terms of Supreme Court opinions,  for each opinion discussed, that discussion is likely to raise other tangential issues on which the justices will want to draw upon past opinions.  Furthermore, for each opinion that applies to and is cited by the current opinion, there is often a case/opinion that needs to be discussed in terms of why it does not apply. Justices often clarify not only those legal principles that apply, but often those that do not. When multiple lines of legal reasoning are drawn upon to justify a decision, often more are discussed in terms of their lack of applicability.
